---
title: "PA1_template"
author: "PNS"
date: "September 16, 2015"
output: html_document
---


## Loading and preprocessing the data

```{r Include packages and datasets, echo=TRUE}
library(knitr)
library(datasets) 
library(ggplot2)
library (scales)
library(Hmisc)

if(!file.exists("activity.csv"))
{unzip("activity.zip")}

activity_dfm <- read.csv("activity.csv")

```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day

2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day


3. Calculate and report the mean and median of the total number of steps taken per day

```{r echo=TRUE}
stepsperday <-aggregate(steps ~ date,activity_dfm,sum)

hist(stepsperday$steps, main = ("Total number of steps taken each day"),col="green", xlab="Total steps per day")

steps_mean <-mean(stepsperday$steps)

paste("Steps Mean is :", steps_mean)

steps_median <-median(stepsperday$steps)

paste("Steps Median is :", steps_median)

```

## What is the average daily activity pattern?

1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r echo=TRUE}

ave_stepsper_interval <- aggregate(x=list(meansteps=activity_dfm$steps),by=list(interval=activity_dfm$interval), FUN=mean, na.rm=TRUE)

maxsteps <- which.max(ave_stepsper_interval$meansteps)
paste("Max number of steps per day is :", maxsteps)

maxtime <- gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", ave_stepsper_interval[maxsteps,'interval'])
paste("Max interval is at  :", maxtime)

ggplot(data=ave_stepsper_interval, aes(x=interval, y=meansteps)) + geom_line() + xlab("Time Interval = 5mins") + ylab("Average number of steps taken") 
```

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r}

num_NA <- length(which(is.na(activity_dfm$steps)))

newactivity_dfm <- activity_dfm

newactivity_dfm$steps <- impute(newactivity_dfm$steps, fun=mean)

newstepsperday <- tapply(newactivity_dfm$steps, newactivity_dfm$date, sum)

qplot(newstepsperday, xlab='New total steps per day', ylab='Total steps per day')

newstepsperday_mean <- mean(newstepsperday)
paste("New Steps Mean is :", newstepsperday_mean)

newstepsperday_median <- median(newstepsperday)
paste("New Steps Median is :", newstepsperday_median)


```

## Are there differences in activity patterns between weekdays and weekends?

```{r}
newactivity_dfm$dateType <-  ifelse(as.POSIXlt(newactivity_dfm$date)$wday %in% c(0,6), 'weekend', 'weekday')

newave_stepsper_interval <- aggregate(steps ~ interval + dateType, data=newactivity_dfm, mean)

ggplot(newave_stepsper_interval, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("Time Interval = 5mins") + 
    ylab("Average number of steps taken")
```

